### 2015-10-27 article/ms.tex
Cleaned up a lot of the model text

\documentclass[useAMS,usenatbib]{mn2e}

\usepackage{amsmath,amsfonts,amssymb,amsbsy}
\usepackage{gitinfo2}

\newcommand\article{\textit{Article}}
\newcommand\tc{\textit{The Cannon}}

\newcommand\K{$\mathbf{K}$}
\newcommand\lv{\mathbf{\boldsymbol\ell_n}}
\newcommand\cv{{\boldsymbol\theta}_\lambda}
\newcommand\wv{\mathbf{w_n}}
\newcommand\ssq{s_\lambda^2}
\newcommand\given{|}
\DeclareMathOperator*{\argmax}{\arg\!\max}

\newcommand\teff{$T_{\rm eff}$}
\newcommand\logg{$\log{g}$}
\newcommand\mh{${\rm [M/H]}$}
\newcommand{\set}[1]{$\{#1\}$}
\newcommand{\logeps}[1]{\log_{\epsilon}{\rm #1}}
\newcommand\ew{$\mathcal{W}$}
\newcommand\psiv{\mathbf{\boldsymbol{\psi}_m}}

\title[Detailed Chemical Abundances with \tc{}]{Detailed Chemical Abundances with \tc{}}

% Some arrangement of the following authors:
\author[Casey et al.]{A.~R.~Casey$^1$, M.~Ness$^2$, G.~Gilmore$^1$,
    D.~W.~Hogg$^{2,3,4}$, H.~W.~Rix$^2$ \\ 
$^1$Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge
    CB3 0HA, UK\\
$^2$Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg,
    Germany\\
$^3$Center for Cosmology and Particle Physics, Department of Physics, New York
    University, 4 Washington Pl., room 424, New York, \\
    NY, 10003, USA\\
$^4$Center for Data Science, New York University, 726 Broadway, 7th Floor,
    New York, NY 10003, USA}
\begin{document}

\date{Accepted 2015 XX XX. Received 2015 XX XX; in original form 2015 XX XX}

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2015}

\maketitle

\label{firstpage}

\begin{abstract}
More than 15 million stellar spectra will be acquired by ground-based surveys in
the coming decade. The detailed chemical abundance analysis of these data poses
a significant challenge. \tc{} provides a purely data-driven
approach to stellar parameter (label) determination. Cross-validation tests
demonstrate that the precision achievable is comparable to existing methods that
rely on model atmospheres and spectral synthesis, but \tc{} is $\sim$300,000 times
faster, making it perfectly applicable for surveys. Including individual abundances by extending \tc{} dimensionality can create enormous model difficulties, severely limiting the interpretability of the results. Here
we describe an alternative approach to including individual element abundances
in \tc{}. Using modest simplifications of the physics of spectral line formation,
we demonstrate precisely \textit{why} an approach like \tc{} works so effectively,
and provide a physically-motivated model that describes flux values, given the
stellar parameters. The inclusion of individual chemical abundances does not
impact on model flexibility, allowing for simultaneous training on detailed
chemical abundances (which are formed through well-understood interactions), as well as arbitrary labels where the impact on the spectra
is unknown or poorly understood (e.g., masses, ages). Crucially, we provide a direct link
between data-driven models and theory of model atmospheres and spectral line formation. We demonstrate that
detailed chemical abundances can be measured with a precision of 0.0X-0.0Y~dex
in Z APOGEE spectra on a modern CPU in just one minute. We present abundance
measurements of 15 elements in Z APOGEE spectra from DR12.
\end{abstract}

\begin{keywords}
\end{keywords}

\section{Introduction}
\label{sec:introduction}

Ground-based stellar spectroscopic surveys have enormous scientific potential.
  Key absorption features in stellar spectra permit inferences on the internal 
structure of stars and their detailed chemical composition.  Given a sufficiently
large sample of stars, these inferences can exclude, guide, or constrain models 
over a large range of physical scales: the formation of planets, stars and 
clusters, supernovae yields and nucleosynthetic pathways, as well as the large 
scale evolution of the Milky Way.  For these reasons, an increasing number of 
stellar spectroscopic surveys are being undertaken, each 
increasing in sample size \citep{gaia_eso, galah, lamost, apogee, weave, 4most}.

We propose the spectroscopic data have outgrown the methods used to analyse
them. There exists an inordinate amount of high quality (high-resolution, high S/N)
spectra, far more than what was available just a decade ago, however the techniques 
employed to extract information from these data have improved only fractionally. 
A subjective (human-driven) analysis was manageable given the data volume, and 
for this reason high-resolution ($\mathcal{R} > 20,000$) stellar spectroscopy has 
colloquially been considered as a `dark art'. Analyses were carefully performed 
by an expert spectroscopist, who themselves were trained by an expert in exactly
where the stellar continuum ought to be placed, which atomic lines are 
considered `clean' or `dirty' (blended), as well as which analysis decisions are
acceptable, and which are not. This work is equally impressive as it is 
subjective: the best spectroscopists host an encyclopaedic knowledge of atomic
absorption lines with decades of experience, yet in blind-tests their decisions 
of where to place the continuum quickly becomes the most dominant source of 
systematic error \citep{who, who}.

Classical spectroscopic analyses are human time-intensive, and were unable to
scale to the current data volume. The Gaia-ESO Survey has been particularly 
influential in minimising the human resources required for analysis: all analysis 
codes of FGK stars employed in Gaia-ESO are completely automatised.  However
a computing challenge has arisen: some analysis codes take up to one hour (single
core) to determine stellar parameters (\teff, \logg, \mh) for a single star. Given the 
expected data growth in the next decade, this poses a major computational error.

These analyses also have strict limitations in validity, and a range of options
are available.  Many analysis options do not have straightforward answers. The 
community standard is to perform chemical abundance analyses using a 
one-dimensional description of a steady-state model photosphere with the
assumption of local thermal equilibrium (LTE). There are a number of 
photospheric models available from different groups.  Detailed information about
the species (atomic and molecular) opacities is required, and partition functions
are usually approximated from polynomials published decades earlier. Atomic 
data (e.g., oscillator strengths) must be well-known, and the treatment of 
natural, thermal and pressure broadening must be appropriately included,
with careful consideration to the impact on perturber collisions. Results are 
further sensitive to the manner in which the equation of state was solved. 
Simply put, there is a lot of physics to consider.

In contrast, \tc{} \citep{Ness15a} ignores everything we know about physics.
\tc{} falls into the regime of supervised machine learning, and is $\sim10^5$
times faster than standard spectroscopic analysis codes. The results published 
using \tc{} thus far are extremely impressive. There is no knowledge of 
stars, no input on interactions between electromagnetic fields and particles, nothing considering
spectral line formation, and no input of atomic data, opacities, partition functions 
or depth-dependent photospheric quantities. Only a relatively small training set
and a polynomial model of stellar parameter labels is required in order to 
reasonably approximate expensive spectral syntheses. However because 
\tc{} is purely data-driven, there is a missing link to physics. Most crucially,
because \tc{} works in label space, including individual chemical abundances
is a non-trivial exercise. Massive increases in the label complexity would be
required to account for known covariances between parameters and elements,
which increases the likelihood of overfitting and limits robustness and interpretability.
Given the high quality spectra currently being collected by ground-based surveys,
extending functionality of \tc\ to include individual chemical abundances -- whilst
maintaining interpretability -- has the potential to revolutionise the way we 
treat stellar spectroscopic data.

This \article{} is organised as follows. In Section \ref{sec:model} we briefly describe
the existing foundations for \tc{} and demonstrate from a theoretical perspective why
the model flexibility is sufficient to accurately represent spectral fluxes. This 
introduction is crucial to outlining how we later include individual atomic lines within 
an existing \textit{Cannon} model. In Section \ref{sec:apogee} we apply a detailed
chemical abundance \textit{Cannon} to the APOGEE Data Release (DR) 12 spectra.
We discuss the suitability and limitations of the method in Section \ref{sec:discussion}.
In Section \ref{sec:conclusion} we conclude with a summary of our findings.

\section{Model}
\label{sec:model}

\tc{} provides a purely data-driven approach to stellar label determination. The
spectral model describes the flux $f_{n\lambda}$ (for star $n$) at a pixel with rest-wavelength $\lambda$ to be a function
of a coefficient vector $\cv$, and a label vector $\lv$:

\begin{equation}
    f_{n\lambda} = g(\lv\given\cv) + {\rm noise}.
    \label{eq:cannon}
\end{equation}

This form is flexible: the label vector $\lv$ may be linear in just a few common
labels (e.g., \teff, \logg, \mh\footnote{\citet{Ness15a} use [Fe/H] in $\lv$, 
but for the sake of distinguishing between overall metallicity and individual 
abundances, throughout this \article{} we opt for \mh\ to describe overall 
metallicity.}) or a complex combination of all $K$ available labels, some of which
may have an unknown or non-trivial relationship with spectral fluxes \citep[e.g.,
ages and masses, see][]{Nissen15, Ness15b}. The underlying principle is that the 
flux for any pixel from any star is predictable with a simple model, given the
labels. Indeed, \citet{Ness15a} employ a quadratic-in-labels spectral model 
using stellar parameters ($T_{\rm eff}$, $\log{g}$, [M/H]) and demonstrate 
through cross-validation that these labels can be measured with a precision of 
95~K in $T_{\rm eff}$, 0.24 in $\log{g}$, and 0.08 in [M/H] with negligible biases.

The noise in Equation \ref{eq:cannon} can be taken as a quadrature combination
of the variance in pixel flux $\sigma_{n\lambda}^2$ and the intrinsic variance 
of the spectral model at each wavelength $s_\lambda^2$, which, like $\cv$, is
not known before the model is trained (see below).

The flux $f_{n\lambda}$ can then be written as a linear function of the label 
vector, leading to the single-pixel log-likelihood function

\begin{equation}
\ln{p}\left(f_{n\lambda}\given\cv^\intercal,\lv, s_{\lambda}\right) = -\frac{1}{2}\frac{\left[f_{n\lambda} - \cv^{\intercal}\cdot\lv\right]^2}{s_\lambda^2 + \sigma_{n\lambda}^2} -\frac{1}{2}\ln\left(s_{\lambda}^2 + \sigma_{n\lambda}^2\right).
\label{eq:pixel-log-likelihood}
\end{equation}

The parameters \set{\cv,s_\lambda^2} for each pixel with wavelength $\lambda$
need to be inferred. This is accomplished given some sample of stars where 
the labels are known with high fidelity (i.e., the training set). Here we will 
optimise the negative log-likelihood (Equation \ref{eq:pixel-log-likelihood}),
which we later confirm to be sufficient for our purposes. This is the training 
step of \tc{}, where the parameters \set{\cv,s_\lambda^2} are optimised 
for a single pixel at $\lambda$ by considering all stars $N$:

\begin{equation}
\cv,s_\lambda \leftarrow \substack{\mbox{arg\,max}\\{\{\cv , s_\lambda\}}} \sum_{n=1}^N \ln{p}\left(f_{n\lambda}\given\cv^\intercal,\lv,s_\lambda^2\right)
\end{equation}

Once the pixel coefficients and variance \set{\cv, s_\lambda^2} are known,
\tc{} provides an approximate \textit{generative model} for spectral fluxes. 
Through a cheap linear product of $\cv^\intercal$ and $\lv$, \tc{} is 
effectively (and efficiently) approximating spectral synthesis, an otherwise
extremely expensive endeavour. This provides a number of advantages,
the most straightforward of which is the ability to efficiently determine $\lv$
(i.e., the stellar parameters) for a new star outside the training set. We refer
the reader to \citet{Ness15a} for a more detailed description of \tc{}.

It is not straightforward to add individual chemical abundances as labels and
expect equally impressive results.  Some chemical abundances only affect few
pixels of the spectrum. Other species (element and ionisation stage) only
experience stimulated emission at wavelengths that are frequently blended
with other species, inducing covariance between elemental abundances.
And all transitions will likely have linear (or higher-order) relationships with
stellar parameters.  Addressing these issues by increasing the complexity 
of the label vector introduces additional issues (e.g., over-fitting, interpretability).
Although the existing \textit{Cannon} framework can be extended to include
individual chemical abundances, they cannot be introduced by extensions 
of the label vector without considering these additional issues.

\subsection{Using the behaviour of spectral lines to understand why \tc\ works}

Before we extend \tc{} to include individual chemical abundances, it is a useful
 exercise to understand \textit{why} a simple spectral model can 
approximate synthesis calculations with comparable fidelity, and thus 
\textit{how} this framework can be extended to include individual abundances. At
present the applications of \tc{} have been restricted to stellar spectra with 
resolving powers of $\mathcal{R} = 22,500$ \citep[APOGEE;][]{Ness15a} or less 
\citep[e.g., LAMOST;][]{Ho15}. In these regimes, the shapes of unsaturated 
spectral lines are well-represented by Gaussian profiles. This leads to several 
crucial insights, which we describe below.

In stellar spectroscopy we are frequently presented with the idea of the 
equivalent width (EW; $\mathcal{W}$) of a spectral line. Consider a single 
(unblended) atomic line at rest wavelength $\lambda_r$, where the flux values 
surrounding $\lambda_r$ ($f_{n\lambda_r-{\delta}},f_{n\lambda_{r+\delta}}$) are 
well-approximated by a \textit{Cannon} spectral model. Because the absorption 
line is well-represented by a Gaussian profile at this spectral resolution, the 
flux at $\lambda_r$ is related to the EW of the line in star $n$ ($\mathcal{W}_n$):

\begin{equation}
\mathcal{W}_n = \sigma{}\sqrt{2\pi}(1 - f_{n\lambda_r}).
\label{eq:ew-flux}
\end{equation}

Here $\sigma$ is the standard deviation of the profile. For typical FGK-type
stars observed with $\mathcal{R} \lesssim 20,000$, $\sigma$ will be dominated
by the line spread function (LSF) of the spectrograph. Because \tc{} is best suited
to large spectroscopic surveys, we can presume that the LSF is well-known, or
at least reasonably constant for the training set, perhaps approximated by the 
quoted spectral resolution $\mathcal{R}$. In Section \ref{sec:something} we use
the knowledge of the spectrograph LSF, but for explanatory purposes here we will
assume it is constant for all $n$ stars surrounding the profile at $\lambda_r$.

If the profile width is constant, then it is straightforward to see by integrating
\tc\ model the entire profile, \tc\ is implicitly modelling the \ew\ of an atomic line
as a function of stellar labels, and modelling spectral flux through a per-pixel
scaling of $\sigma_\lambda$. Indeed, the success of \tc\ implies that the 
strength \ew\ of an atomic line could be well-approximated by a quadratic
model of a label vector containing the stellar parameters, where \mh\ is 
actually a proxy for the abundance of the atomic line in question. If such a
\textit{line strength model} could be constructed and appropriately
included within the existing \textit{Cannon} framework, it would permit the 
determination of individual chemical abundances, whilst maintaining the 
flexibility in training on labels with poorly-known relationships on spectral 
fluxes (e.g., ages).

We note that Equation \ref{eq:ew-flux} does not imply that \tc{} is limited in
applicability to low-resolution ($\mathcal{R} \lesssim 20,000$) spectra. Quite 
the opposite. \tc{} is sufficiently flexible that in extremely high-resolution 
spectra, where absorption lines are more closely represented by Lorentzian or
Hertzting functions, the flux values $f_{n\lambda}$ in the wings of a strong 
line are still well-modelled by \tc{} (as judged by $\ssq$ and performance in
cross-validation tests). Indeed, the line strength \ew\ would still be 
predictable with high-fidelity, implying that the coefficients $\cv$ for pixels
in the wings just scale in a different way.

The crucial inference from Equation \ref{eq:ew-flux} is that knowing the LSF
allows us to directly translate \ew\ to $f_{n\lambda}$, thereby accounting for 
the flux contribution from a single atomic line across all neighbouring pixels.
If the spectral resolving power were sufficiently high such that the absorption
profiles appeared Lorentzian, some prior information on the shape $\gamma$
may be required to translate \ew\ to $f_{n\lambda}$. Importantly, all future 
stellar spectroscopic surveys are currently planning for spectral resolutions
$\mathcal{R} \lesssim 20,000$, in the perfect Gaussian-profile regime for this
extension of \tc{}.

To demonstrate that \tc\ is implicitly modelling the EW of an atomic line as
a function of stellar parameters and the abundance of the given element,
here we employ the quadratic-in-labels model from \citet{Ness2015a} to
model the strength of a single atomic line. We used the NNN stars from the
training set used by \citet{Ness2015a} and calculated the abundance of a
Si 1 line at N \AA\ for thirty different EWs uniformly sampled between 5~m\AA\
and 120~m\AA. These stars cover the main sequence and red giant branch
(RGB), and the overall metallicity spans three decades from -2.5 to +0.5 dex.
The abundances were calculated using MARCS model atmospheres \citep{marcs}
and a modified Cython version of MOOG \citep[][2014 version]{Sneden1973}
which includes ABO broadening of atomic lines \citep{ABO} and several 
improvements to the treatment of hydrogen opacities. The atomic data
employed is listed in Table \ref{tab:atomic-data}. We describe the 
\textit{line strength model} that relates the line strength \ew\ for the $m$-th
atomic line in star $n$ as,

\begin{equation}
\mathcal{W}_{n,m} \approx \psiv^{\intercal}\cdot\lv
\label{eq:line-strength-model}
\end{equation}

\noindent{}where the label vector 
$\lv = \{T_{\rm eff},\log{g},\log_{\epsilon}{\rm Si}\}$\footnote{Recall
${\rm [Si/Fe]} \equiv {\rm [Si/H]} - {\rm [Fe/H]} \equiv \log_{\epsilon}{\rm Si} 
- \log_{\epsilon}{\rm Si}_\odot - {\rm [Fe/H]}$, where we adopt the \citet{Asplund2009}
solar abundances.} and the coefficients of the vector $\psiv$ for a given 
atomic line need to be solved for. Combining Equations \ref{eq:ew-flux} and
\ref{eq:line-strength-model} yields the approximate flux $f_{n,\lambda_r}$ that 
atomic line $m$ contributes to star $n$ at wavelength $\lambda_r$, which we
describe as $h(\lv|\psiv,\sigma)$:

\begin{equation}
f_{n,\lambda_r} \approx h(\lv|\psiv,\sigma,\lambda) \equiv 1 - \frac{\psiv^{\intercal}\cdot\lv}{\sigma\sqrt{2\pi}}
\end{equation}

We calculate $\psiv$ given the stellar parameters of the training set and the
calculated $\logeps{(Si~1)}$ abundances from MOOG, calculated at many \ew\ values.
The difference between the MOOG-calculated abundance and our line strength model
is shown in the second panel of Figure \ref{fig:atomic-line-model-example}.
Negligible bias is present, and the standard deviation is 0.01~dex, demonstrating
that a quadratic-in-labels model can relate line strength \ew\ and abundance
for all types of stars in the training set. This is further shown in the third panel of 
Figure \ref{fig:atomic-line-model-example},
which shows the spectrum surrounding the Si 1 line for a random star\footnote{The
star is truly chosen at random, see Appendix \ref{appendix:reproducibility} for steps
to reproduce this research.} in the training set. We calculated the corresponding
line strength \ew\ given the stellar parameters and reported Si 1 abundance, 
adopted a spectral resolution of $\mathcal{R} = 22,500$ and produced a model 
spectrum of this atomic line following Equation \ref{ew-flux}.

We note that some of the 0.01~dex scatter between our line strength model and 
MOOG (second panel of Figure \ref{fig:atomic-line-model-example}) is likely due to 
the tolerance employed by MOOG's optimisation algorithm to calculate a line 
abundance, given an EW.

% Show a figure with different EW vs abundances, colored by teff/logg
% Then show the expected EW vs synthesis EW, and show the residual
% Show a portion of the spectrum where we translate a modelled EW into a spectrum
% by using some knowledge of the LSF

% h relates W to flux values.
% W

% Discuss using the Cannon model from Ness et al to model the EW as a func. of stellar params.

\subsection{Using the behaviour of spectral lines to guide model complexity}

Knowing that a simple model can relate the a single line abundance 
$\log_{\epsilon}{\rm X}$ to the equivalent width \ew\, given the stellar
parameters, it is intriguing to ask whether model atmosphere relations and
the theory of spectral line formation can guide the level of model complexity
required.

% Show an example? The example could have a figure showing some spectral fluxes for a clean line, which is modelled by the cannon (with an example model spectrum) and then show the same model being used to train on EW.
%An example of this is shown in Figure \ref{fig:cannon-ew-relation}, where a quadratic-in-labels \textit{Cannon} model is used to model spectral fluxes for X stars across the HR diagram. The data have been trained

% How does EW of a line change w.r.t. stellar parameters
\textbf{DESCRIPTION OF HOW EW FOR A SINGLE LINE WILL CHANGE WITH RESPECT TO STELLAR PARAMETERS AND ABUNDANCE FOR A NEUTRAL LINE IN A GRAY ATMOSPHERE.}

\subsection{Extending \tc{} to include individual atomic lines}

We now have a functional form of a model that relates $\mathcal{W}$ to $\log_\epsilon\left({\textrm X}\right)$ (or a scaled-solar [X/H] value), given the stellar parameters, which themselves are part of the set of $K$ labels. It is interesting to note that Equation \ref{eq:ew-logx-model} is not dissimilar to the quadratic-in-labels model used by \citet{Ness15a} for APOGEE data. With only the prior knowledge that \textit{a simple model of the labels can predict spectral fluxes}, we have derived the appropriate functional form for a single absorption line. The model $\mathcal{W}\left({\rm [X/H]}, T, logg\right)$ simply approximates the curve-of-growth for a single line, and the coefficients still need to be calculated. The practicalities of determining these coefficients is outlined in Section \ref{sec:somewhere}. For the purposes of introducing the model, we can assume that they are easily calculable.

Consider multiple atomic lines, where the coefficients to translate $\mathcal{W}$ and [X/H] are known. Given some knowledge of the line spread function, the EWs $\mathcal{W}$ to flux values $f_{n\lambda}$ can be translated such that:

\begin{equation}
h = X
\label{eq:h-model}
\end{equation}

Here we implicitly assume the mutual opacity contributions of nearby weak lines will be small. 
At this point it is worthwhile to ask whether Equation \ref{eq:h-model} can be scaled sufficiently to $M$ atomic lines in order to produce an entire spectral model for $f_{n\lambda}$, without the contribution of $g\left(\lv\given\cv\right)$. There would be a number of immediate issues with that approach: a huge number of additional coefficients would be required, hidden lines could not be encapsulated accurately by the model, strong lines would be extremely poorly modelled by the fit, no longer linear (slow), but most importantly the pixels would lose flexibility in training on arbitrary labels.

If lines are blended, one could consider modelling many more atomic transitions.
However as the number grows, this situation becomes unnecessarily complex:
often blended lines exist and we do not care about their abundance, only their
inclusion (nuisance or dirty lines). Other features may be a complex combination
of thousands of molecular features.

Here we introduce a method of including individual atomic transitions within an existing \textit{Cannon} model.

The spectral model becomes a multiplicative function of $g$ and $h$ with some
unknown noise.

\begin{equation}
    f_{n\lambda} = g(\lv\given\cv) \cdot h(\lv,\wv\given\psi_n) +  {\rm noise}.
\end{equation}

This inclusion combines the existing flexibility in the model, allowing for training on arbitrary labels, and uses individual absorption lines for which some information is known (species, atomic data, etc) in order to trace the chemical abundances of individual elements.

The vector $\wv$ can be calculated directly from spectral synthesis, or using data.
In practice, given some rest-frame normalised spectra and some knowledge of the
line spread function, $h(\lv,\wv\given{}X)$ can be calculated before training \tc.

%Utilising a simple polynomial model to predict spectral fluxes is a distinctive
%method from parallel efforts in radiative transfer of realistic (3D) stellar 
%photospheres. Calculations of this nature are extremely expensive. Indeed, even
%with simplifications of these models (e.g., interpolated 1D photospheres),
%calculating synthetic spectra requires knowledge of species opacities, detailed 
%atomic and molecular information for all contributing elements, solving for
%the equation of state, and iteratively performing expensive radiative transfer
%calculations. Given the
%simplicity of a quadratic-in-labels spectral model with \tc{}, it is remarkable
%that these complex calculations can be simplified so greatly, whilst maintaining
%the typical measurement fidelity reported for high-resolution spectroscopic studies based on
%\textit{curve-of-growth} analyses.

% Figure of relating EW to [X/H] with some simple model compared to analytic value.

The compromise is to use TC and `clean' lines that we care about in concert.

% existing simple model. replaces expensive synthesis, involving model atmosphere
% generation, t-tau relations, opacities, equation of state, integrated line-of-sight
% and radiative transfer, etc.

% before extending this model to include individual abundances, it is worthwhile
% demonstrating (from a physics perspective) *why* the cannon works so well.

% APOGEE, R of 20,000, 
% LAMOST, higher Rs
% super high-resolution also works.

\section{APOGEE}

% All about APOGEE Survey and existing data.

% Apply a model using stellar parameters and abundances from X stars in clusters
% Typical S/N ratio and quality constraints employed
% Any abundance flags ignored?

% Train a model based on stellar parameters, include XYZ lines (Tabular data)

\section{Discussion}

% Comparison in EW model to log(X)
% Std deviation in this model.

% Training directly from data, or training from physical models?
% Inferences that can be made from models (when data are lacking)

% Requirements about knowing the LSF (e.g., a one-parameter LSF is just as good as TC)
% Blended lines

% Limitations: what about missing abundance data?

\section{Conclusions}

%

\section*{Acknowledgments}
This research was fostered by discussions at the \textit{New Milky Way}
conference hosted by the Munich Institute for Astro- and Particle Physics, and
the \textit{Stellar Streams in the Local Universe} conference hosted at Ringberg
Castle by the Max-Planck-Institut F\"ur Astronomie. This research made use of 
Astropy, a community-developed core Python package for Astronomy \citep{astropy}.
The authors are thankful for the following free online services which help make
science reproducibility easier: GitHub, Travis Continuous Integration, and Zenodo.

\begin{thebibliography}{99}

\bibitem[Asplund et al.(2009)]{Asplund2009} Asplund, M., Grevesse, N., Sauval, A.~J., \& Scott, P.\ 2009, \araa, 47, 481

\bibitem[Astropy Collaboration et al.(2013)]{astropy} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., et al.\ 2013, \aap, 558, AA33

\bibitem[Ho et al.(2015)]{Ho15} Ho, A., et al., in preparation
\bibitem[Ness et al.(2015a)]{Ness15a} Ness, M., Hogg, D.~W., Rix, H.-W., Ho, A.~Y.~Q., \& Zasowski, G.\ 2015, \apj, 808, 16

\bibitem[Ness et al.(2015b)]{Ness15b} Ness, M., et al., in preparation

\bibitem[Nissen(2015)]{Nissen15} Nissen, P.~E.\ 2015, \aap, 579, A52

\end{thebibliography}

\label{lastpage}

\appendix
\section{Reproducing this work}
\label{appendix:reproducibility}

This \article{} has been compiled from the \texttt{git} repository (abbreviated commit hash \texttt{\gitAbbrevHash} on \texttt{\gitBranch} branch) hosted online by GitHub at \texttt{github.com/andycasey/fireworks}. The data required to reproduce this analysis is persistently hosted through Zenodo \citep{DATA}. The complete analysis, figures and tables included in this \article{} can be reproduced in full using the following commands in a modern Unix-based terminal:
\vspace{0.5em}

\noindent\texttt{[reproducibility commands]} 
\vspace{0.5em} \\
If the required tools are not available on the readers system (e.g., \texttt{git}), this research can also be produced using the public Docker image \texttt{[docker image reference]}.

\end{document}